---
title: "Cross-Cultural Quantitative Research: Trust and Leadership"
author: "Monika Bagociute"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    code_folding: hide
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '3'
---


## Research overview

This study examines how different approaches to vaccine allocation
relate to trust in political leadership across participants from four
English-speaking countries.

The analysis demonstrates an end-to-end quantitative research workflow,
including data preparation, quality screening, survey-based measurement,
statistical modelling, effect estimation, exploratory analysis, and
sensitivity analysis.

### Research question

Does the framing of vaccine allocation influence reported trust in
political leadership, and does this relationship vary across participant
characteristics and countries?

### Analytical approach

The primary analysis estimates the relationship between experimental
condition and trust while accounting for relevant participant
characteristics and country-level variation. Exploratory analyses
investigate whether the relationship differs across countries and
participant characteristics.

### Tools

- R
- R Markdown
- Linear and mixed-effects modelling
- Estimated marginal means
- Confidence intervals
- Effect sizes
- Data visualisation

## 1. Reproducible Analysis Setup

```{r setup, include=FALSE}
kable <- function(data) {
  knitr::kable(data, format = "html", digits = 2) %>%
        kable_styling(bootstrap_options = "striped", full_width = F,
                      position = "left")
}
```

```{r, message=F}
library(dplyr)
library(tidyr)
library(kableExtra)
library(ggplot2)
library(lme4)
library(lmerTest)
library(emmeans)
library(summarytools)
library(rstatix)
library(effectsize)
library(coin)

```

```{r}
set.seed(828)
```

#### Plot Setup

```{r}
theme_set(theme_classic(base_size = 12) +
          theme(axis.title = element_text(size = 15, face='bold'),
                axis.title.x = element_text(margin = margin(6, 0, 0, 0)),
                axis.title.y = element_text(margin = margin(0, 6, 0, 0)),
                axis.ticks.y = element_line(color='#4d4d4e'),
                axis.ticks.x = element_blank(),
                strip.background = element_blank()))
```

## Data Preparation

```{r}
# data_raw <- read.csv("S1 - Data.csv")

data_raw <- read.csv(
  file.path("..", "Data", "raw", "study1_data.csv")
)
```

N = `r  nrow(data_raw)`

### Recode variables

```{r}
data_raw %>%
  
  mutate(vnat_b = gsub(".*ensure that there are enough vaccines for people.*", "Keep",
                       gsub(".*ensure that people in developing countries have access to vaccines.*",
                            "Distribute", vnat_b)),
         
         gender_rec = if_else(gender %in% c("Woman", "Man"), gender, "Other"),
         gender_rec = factor(gender_rec, levels = c("Woman", "Man", "Other")),
         
         race_rec = gsub("Australian", "White", race),
         race_rec = if_else(race_rec == "White", "White", "Other"),
         race_rec = factor(race_rec, levels = c("White", "Other")),
         
         country = factor(country, levels = c("United States", "Australia", 
                                               "Canada", "United Kingdom")),
         
         edu_rec = recode(education, "Some elementary school / primary school" = "1",
                             "Completed elementary school / primary school" = "2", 
                             "Some high school / secondary school" = "3", 
                             "Completed high school / secondary school" = "4", 
                             "Some college / undergraduate degree" = "5", 
                             "Completed college / undergraduate degree" = "6", 
                             "Some advanced postgraduate degree" = "7", 
                             "Completed advanced postgraduate degree " = "8",
                             "Completed advanced postgraduate degree" = "8"),
         edu_rec = as.numeric(edu_rec),
         
         cond = factor(cond, c("Redistribution", "Nationalism"))) -> data_raw
```


#### Create Composite measure of trust

```{r}
data_raw$trust_avg <- (data_raw$trust_1+data_raw$trust_2)/2
```
#### Apply effect coding to experimental condition

```{r}
data_raw$cond_rec <- recode(data_raw$cond, "Nationalism" = -0.5, "Redistribution" = 0.5)
```

## 3. Data Quality and Exclusions

```{r}
data_raw %>%
  mutate(failed_att1 = att_check1 != "TikTok",
         failed_att2 = att_check2 != "Roses can suffer with pests like aphids and blackfly",
         failed_comp = comp_check != "How much I trusted the mayor") -> data

data %>%
  select(sub_n, trust_1, trust_2, gender, age, race, 
         country, education, ses2, politics, religious) %>%
  mutate(n_missed_key = rowSums(is.na(.)), 
         missed_half = n_missed_key > 10/2) %>%
  select(sub_n, n_missed_key, missed_half) %>%
  right_join(data, by='sub_n') -> data
```

Removing:

- `r sum(data$duplicate_IP + data$duplicate_ID > 0)` participated more than once

```{r}
data <- filter(data, duplicate_IP == 0, duplicate_ID == 0)
```

- `r sum(data$failed_att1 + data$failed_att2 > 0)` failed 1+ attention check

```{r}
data <- filter(data, failed_att1 == 0, failed_att2 == 0)
```

- `r sum(is.na(data$country) | (data$country != data$survey_c))` in different country  

```{r}
data <- filter(data, survey_c == country)
```

- `r sum(data$missed_half != 0)` responded to less than half of key questions  

```{r}
data <- filter(data, missed_half == 0)
```

- `r sum(data$failed_comp != 0)` failed comprehension check

```{r}
data <- filter(data, failed_comp == 0)
```

Post-exclusions N = `r nrow(data)`

`r kable(count(data, country))`

## 4. Sample Characteristics

N women: `r sum(data$gender == "Woman")`

Mean age: `r round(mean(data$age, na.rm = TRUE), 2)`

Center covariates

```{r}
data$politics_raw <- data$politics
data$support_raw <- data$support
data <- mutate_at(data, c("age", "edu_rec", "ses2", "politics",
           "religious", "support"), ~scale(.x, center = TRUE, scale=FALSE))

saveRDS(
  data,
  "../Data/processed/cleaned_data_s1.rds"
)
```

Set contrasts

```{r}
options(contrasts = c("contr.treatment", "contr.poly"))
```

## 5. Primary Analyses

Country random intercept:

```{r}
mod <- lmer(trust_avg ~ gender_rec + age + race_rec 
          + edu_rec + ses2 + politics + religious 
          + support + cond_rec 
          + (1|country), data = data)
```

```{r}
kable(summary(mod)$coefficients)
```

#### Estimate a simplified model without the country random intercept

```{r}
mod <- lm(trust_avg ~ gender_rec + age + race_rec 
          + edu_rec + ses2 + politics + religious 
          + support + cond_rec, data = data)
```

```{r}
kable(summary(mod)$coefficients)
```

```{r}
kable(confint(mod, "cond_rec", level = 0.95))
```

```{r}
kable(emmeans(mod, specs = "cond_rec"))
```

```{r fig.width=3, fig.height=3}
data %>%
  group_by(cond) %>%
  summarise(mean = mean(trust_avg), sd = sd(trust_avg), 
            count = n(), se = sd / sqrt(count), 
            ci = se * qt(1-(0.05/2), count-1))  %>%
  mutate(xpos = row_number()) %>%
  ggplot(aes(x=cond, y = mean, color=cond, fill=cond)) + 
  geom_jitter(data = data, aes(y = trust_avg), width = .2,
              height = 0, alpha = .3, size = 0.75, shape = 16) +
  geom_crossbar(aes(ymin = mean-ci, ymax = mean+ci),
                alpha = .2, linewidth = 0.4, fatten = 1.75, width = .75) +
  scale_fill_manual(values = c("#176f8d", "#158d71")) +
  scale_color_manual(values = c("#176f8d", "#158d71")) +
  xlab('Leader') +
  ylab('Trust in Leader') +
  theme(legend.position = "none") +
  scale_y_continuous(breaks=seq(1, 7, 2), limits=c(0.5, 7.5)) 
```

```{r}
ggsave("S1 - Fig 1.pdf", width = 2.75, height = 2.45)
```

## 6. Exploratory Analyses

### Effect by country:

```{r}
mod <- lm(trust_avg ~ cond_rec * country, data = data)
```

```{r}
kable(summary(mod)$coefficients)
```

```{r fig.width=10, fig.height=3}
data %>%
  group_by(cond, country) %>%
  summarise(mean = mean(trust_avg), sd = sd(trust_avg), 
            count = n(), se = sd / sqrt(count), 
            ci = se * qt(1-(0.05/2), count-1))  %>%
  mutate(xpos = row_number()) %>%
  ggplot(aes(x=cond, y = mean, color=cond, fill=cond)) +
  geom_jitter(data = data, aes(y = trust_avg), width = .2,
              height = 0, alpha = .3, size = 0.75, shape = 16) +
  geom_crossbar(aes(ymin = mean-ci, ymax = mean+ci),
                alpha = .2, linewidth = 0.4, fatten = 1.75, width = .75) +
  scale_fill_manual(values = c("#176f8d", "#158d71")) +
  scale_color_manual(values = c("#176f8d", "#158d71")) +
  xlab('Leader') +
  ylab('Trust in Leader') +
  theme(legend.position = "none") +
  scale_y_continuous(breaks=seq(1, 7, 2), limits=c(0.5, 7.5)) +
  facet_wrap(~country, nrow = 1)
```

```{r}
ggsave("S1 - Fig 2.pdf", width = 8.40, height = 2.64)
```

### Without covariates

```{r}
mod <- lm(trust_avg ~ cond_rec, data = data)
```

```{r}
kable(summary(mod)$coefficients)
```

```{r}
kable(confint(mod, "cond_rec", level = 0.95))
```

### Participant Politics

Nationalism:

```{r}
test <- t.test(filter(data, vnat_b == "Keep", 
                      cond == "Redistribution")$trust_avg,
               filter(data, vnat_b == "Keep", 
                      cond == "Nationalism")$trust_avg)

efsz <- cohens_d(filter(data, vnat_b == "Keep", 
                      cond == "Redistribution")$trust_avg,
               filter(data, vnat_b == "Keep", 
                      cond == "Nationalism")$trust_avg)
```

t(`r round(test$parameter)`) = `r round(test$statistic, 2)`, p = `r round(test$p.value, 3)`, d = `r round(efsz$Cohens_d, 2)`, CI = [`r round(efsz$CI_low, 2)`, `r round(efsz$CI_high, 2)`]

```{r fig.width=3, fig.height=3}
data %>%
  filter(vnat_b == "Keep") %>%
  group_by(cond) %>%
  summarise(mean = mean(trust_avg), sd = sd(trust_avg), 
            count = n(), se = sd / sqrt(count), 
            ci = se * qt(1-(0.05/2), count-1))  %>%
  mutate(xpos = row_number()) %>%
  ggplot(aes(x=cond, y = mean, color=cond, fill=cond)) + 
  geom_jitter(data = data, aes(y = trust_avg), width = .2,
              height = 0, alpha = .3, size = 0.75, shape = 16) +
  geom_crossbar(aes(ymin = mean-ci, ymax = mean+ci),
                alpha = .2, linewidth = 0.4, fatten = 1.75, width = .75) +
  scale_fill_manual(values = c("#176f8d", "#158d71")) +
  scale_color_manual(values = c("#176f8d", "#158d71")) +
  xlab('Leader') +
  ylab('Trust in Leader') +
  theme(legend.position = "none") +
  scale_y_continuous(breaks=seq(1, 7, 2), limits=c(0.5, 7.5))
```

```{r}
ggsave("S1 - Fig 3.pdf", width = 2.75, height = 2.45)
```

Conservatives:

```{r}
test <- t.test(filter(data, politics_raw == 7, 
                      cond == "Redistribution")$trust_avg,
               filter(data, politics_raw == 7, 
                      cond == "Nationalism")$trust_avg)

efsz <- cohens_d(filter(data, politics_raw == 7, 
                      cond == "Redistribution")$trust_avg,
               filter(data, politics_raw == 7, 
                      cond == "Nationalism")$trust_avg)
```

t(`r round(test$parameter)`) = `r round(test$statistic, 2)`, p = `r round(test$p.value, 3)`, d = `r round(efsz$Cohens_d, 2)`, CI = [`r round(efsz$CI_low, 2)`, `r round(efsz$CI_high, 2)`]





